environmental_data_preperation

### 2018-11-12 src/env_data_prep.Rmd
i crashed the environmental_data_preperation notebook... recreated in env_data_prep

reproject ...
There is currently an issue with the projectRaster command in the raster package. Michael Sumners kindly offered a [temporary solution](https://twitter.com/mdsumner/status/1040500830539935744)
```{r}
## use ncdf4 directly
library(ncdf4)
## connection to the file
nc <- nc_open("../data/env/1993_01_salinity.nc")
## desired variable 
v <- ncvar_get(nc, "salinity")
## dimension coordinates
lon <- ncvar_get(nc, "longitude")
lat <- ncvar_get(nc, "latitude")
## a template, one map from the lon, lat, and v[,,1]
template <- raster(list(x = lon, y = lat, z = v[,,1]), crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 ")

## apply template to brick, and set values to v (but flipped in Y because that's how raster works - lat is increasing in the NetCDF, so we go upside down)
diy <- setValues(brick(template), v[, ncol(v):1, ])

## target
prj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs" #note -60 used to be -91. Changed to 'straighten up'.
target <- projectExtent(diy, prj)  ## get a target extent, and we'll set the exact resolution we want
res(target) <- c(25000, 25000)
aea <- projectRaster(diy, target)
plot(aea[[1]]) #the number in brackets refers to the depth number
aea
writeRaster(aea[[1]],'../output/env/aea.tif',overwrite=TRUE,options=c('TFW=YES'))
```

@@

now give a unique ID to each cell
```{r}
unique_cell[unique_cell > 0] <- 1:7768 
writeRaster(unique_cell,'../output/env/glo_unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
uniquecell_poly <- rasterToPolygons(aea[[1]], fun=NULL, n=4, na.rm=TRUE, digits=12, dissolve=FALSE)
writeOGR(uniquecell_poly, dsn = '.', layer = '../output/env/glo_unique_cell_poly', driver = "ESRI Shapefile")

```

@@

shift the dec column you created down by 1
```{r}
library(useful)
naophases <- shift.column(data = naophases, columns = "12", len = 1, up = FALSE)
naophases
write.csv(naophases, file = "../output/env/nao.csv", row.names = FALSE) #for some reason you need to write to a csv and then reload for the next step to work properly...

```

@@

First you need to put the observations (data_aea) into a spatialpointsdataframe.
```{r}
library(sp)
xy2 <- data_aea[ ,c(18,19)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
data_aea_sp <- SpatialPointsDataFrame(coords = xy2, data = data_aea, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
```

and the loop. Note that MLP and SSH only have a single layer and as such they do not have a statement to extract at depth.
```{r}
netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
library (ncdf4)
for (i in 1:no_netcdf){  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:nrow(data_aea_sp)) {  
      print (j)
      de <- data_aea_sp$depthlayerno[[j]]  # a variable for the observation depth layer
      yr <- (data_aea_sp$year[j])  # a variable for the observation year
      mth <- (data_aea_sp$month[j])  # a variable for the observation month
          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){

@@

          }
    } 
  }
}

```

>>>>

reproject ...
There is currently an issue with the projectRaster command in the raster package. Michael Sumners kindly offered a [temporary solution](https://twitter.com/mdsumner/status/1040500830539935744)
```{r}
## use ncdf4 directly
library(ncdf4)
## connection to the file
nc <- nc_open("../data/env/1993_01_salinity.nc")
## desired variable 
v <- ncvar_get(nc, "salinity")
## dimension coordinates
lon <- ncvar_get(nc, "longitude")
lat <- ncvar_get(nc, "latitude")
## a template, one map from the lon, lat, and v[,,1]
template <- raster(list(x = lon, y = lat, z = v[,,1]), crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 ")
## apply template to brick, and set values to v (but flipped in Y because that's how raster works - lat is increasing in the NetCDF, so we go upside down)
diy <- setValues(brick(template), v[, ncol(v):1, ])
## target
prj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs" #note -60 used to be -91. Changed to 'straighten up'.
target <- projectExtent(diy, prj)  ## get a target extent, and we'll set the exact resolution we want
res(target) <- c(25000, 25000)
aea <- projectRaster(diy, target)
plot(aea[[1]]) #the number in brackets refers to the depth number
aea
writeRaster(aea[[1]],'../output/env/aea.tif',overwrite=TRUE,options=c('TFW=YES'))
```

@@

now give a unique ID to each cell
```{r}
unique_cell[unique_cell > 0] <- 1:7768 
writeRaster(unique_cell,'../output/env/glo_unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
uniquecell_poly <- rasterToPolygons(aea[[1]], fun=NULL, n=4, na.rm=TRUE, digits=12, dissolve=FALSE)
writeOGR(uniquecell_poly, dsn = '.', layer = '../output/env/glo_unique_cell_poly', driver = "ESRI Shapefile")
```

@@

shift the dec column you created down by 1
```{r}
library(useful)
naophases <- shift.column(data = naophases, columns = "12", len = 1, up = FALSE)
naophases
write.csv(naophases, file = "../output/env/nao.csv", row.names = FALSE) #for some reason you need to write to a csv and then reload for the next step to work properly...
```

@@

First you need to put the observations (data_aea) into a spatialpointsdataframe.
```{r}
library(sp)
xy <- data_aea[ ,c(18,19)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aea, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
```

```{r}
netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not includedCores
library (ncdf4)
library(sp)
library(rgdal)
library(raster)
for (i in 1:no_netcdf) {  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:nrow(data_aea_sp)) {  
      print (j)
      de <- data_aea_sp$depth[[j]]  # a variable for the observation depth layer
      yr <- (data_aea_sp$year[j])  # a variable for the observation year
      mth <- (data_aea_sp$month[j])  # a variable for the observation month
          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){

@@

          }

    }
}
```

### 2018-11-12 src/environmental_data_preperation.Rmd
i crashed the environmental_data_preperation notebook... recreated in env_data_prep

and the loop. Note that MLP and SSH only have a single layer and as such they do not have a statement to extract at depth.
```{r}
netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
library (ncdf4)
for (i in 1:no_netcdf){  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:nrow(data_aea_sp)) {  
      print (j)
      de <- data_aea_sp$depthlayerno[[j]]  # a variable for the observation depth layer
      yr <- (data_aea_sp$year[j])  # a variable for the observation year
      mth <- (data_aea_sp$month[j])  # a variable for the observation month
          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){

@@

          }
    } 
  }
}

>>>>

and the loop. Note that MLP and SSH only have a single layer and as such they do not have a statement to extract at depth.
```{r}
netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
library (ncdf4)
for (i in 1:no_netcdf){  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:nrow(data_aea_sp)) {  
      de <- data_aea_sp$depthlayerno[[j]]  # a variable for the observation depth layer
      yr <- (data_aea_sp$year[j])  # a variable for the observation year
      mth <- (data_aea_sp$month[j])  # a variable for the observation month
          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){

@@

          }

    }
}

### 2018-11-12 src/raster_sjl.Rmd
i crashed the environmental_data_preperation notebook... recreated in env_data_prep

>>>>

full variables

```{r}
data_aea <- read.csv("../data/observations3.csv", header = TRUE) #change this path to your .csv location
xy <- data_aea[ ,c(4,5)] # This is to tell R where the coordinates are (in column 4 and 5). Note that the column order needs to be longitude, latitude
data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aea, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
```

```{r}

netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not includedCores
library (ncdf4)
library(sp)
library(rgdal)
#no_cores <- detectCores() -1
#core_cluster <- makeCluster(no_cores)
#registerDoParallel(core_cluster)
for (i in 1:no_netcdf) {  
  library(raster)
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:nrow(data_aea_sp)) {  
      print (j)
      de <- data_aea_sp$depth[[j]]  # a variable for the observation depth layer
      yr <- (data_aea_sp$year[j])  # a variable for the observation year
      mth <- (data_aea_sp$month[j])  # a variable for the observation month
          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){

             data_aea_sp$temp_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
              if (is.na(de)){
                data_aea_sp$temp_depth[j] <- NA
              } else  
                data_aea_sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
              data_aea_sp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
              if (is.na(de)){
                data_aea_sp$salinity_depth[j] <- NA
              } else  
                data_aea_sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
              data_aea_sp$chl_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
              if (is.na(de)){
                data_aea_sp$chl_depth[j] <- NA
              } else  
                data_aea_sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
              data_aea_sp$o2_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
              if (is.na(de)){
                data_aea_sp$o2_depth[j] <- NA
              } else  
                data_aea_sp$o2_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
              data_aea_sp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
              data_aea_sp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ])

          }

    }
}

#stopCluster(core_cluster)
```

```{r}
data_aea_sp$mlp_surface
data_aea_sp$salinity_depth
```
